---
title: "Week4 note"
output: html_document
---
# week4
## Editing Text Variables
### fixing character vectors- `tolower(), toupper()`

```{r}
if(!file.exists("./data")) {dir.create("./data")}
fileUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl, destfile = "./data/cameras.csv", method = "curl")
cameraData <- read.csv("./data/cameras.csv")
names(cameraData)
```

```{r}
tolower(names(cameraData))
```
### Fixing chararter vectors- `strsplit()`

- good for automatically spliting variable names
- important parameters:x split
```{r}
splitNames <- strsplit(names(cameraData), "\\.")
splitNames[[5]]  # no change
splitNames[[6]]
```
### quick aside- lists
```{r}
mylist <- list(letters = c("A", "b", "c"),
               numbers = 1:3,
               matrix(1:25, ncol = 5))
head(mylist)
```
```{r}
mylist[1]
```

```{r}
mylist$letters
```

```{r}
mylist[[1]]
```
### sapply()
```{r}
spiltNames[[6]][1]
firstElement <- function(x) {x[1]}
sapply(splitNames, firstElement)
```
### sub()
change _ in rewiews into blank. but just the first _ will be replaced.
`sub("_","",names(reviews),)`
gsub will replace all the _.
`gsub("_", "", "this_is_a_test")`

### grep(), grepl()

```{r}
grep("Alameda", cameraData$intersection)
grep("Alameda", cameraData$intersection, value = TRUE)

grepl("Alameda", cameraData$intersection)
```

### more useful string function

```{r}
library(stringr)
nchar("jeffrey leek")           # count the characters
substr("Jeffrey leek", 1, 7)    # find the 1-7 letters
paste("Jeffrey", "leek")        # paste strings together with a space
paste0("jeffrey", "leek")       # paste strings together
str_trim("jef     ")            # remove white space
```
## Regular Expressions
### Metacharacters
- the start of a line `^`
  `^i think`
- the end of a line `$`
  `morning$`
- list of characters will accept at a given point
  `[Bb][Uu][Ss][Hh]`
- character classes with []
  `^[Ii] am`
- a range of lesters
  `^[0-9][a-zA-z]`
- "^" used at the beginning of a character class ,NOT inthe indicated class
  `[^?.]$`
- "." is uesed to refer to any chararcter
  `9.11`
- "|" or, alternatives
  `flood|fire`
  `flood|earthquake|hurricane|coldfire`
  `^[Gg]ood|[Bb]ad`
  `^([Gg]ood|[Bb]ad)`
- "?" indicated expression is optional
  `[Gg]eogre( [Ww]\. )? [Bb]ush`
- * means "any number, including none of the item" + means "at least one of the item"
  `(.*)`
  `[0-9]+(.*)[0-9]+`
- {} interval quantifiers,specify the minimun and maximum number
  `[Bb]ush( +[^ ]+ +){1,5} debate` 
  1-5 words between bush and debate
- and
  - m,n at least m ,but not more than n matches
  - m means exactly m matches
  - m, means at least m matches
- \1, \2 refer to the matched text
  ` +([a-zA-Z]+) +\1 +`
- the greediness of * can be turened off with the ?
  `^s(.*?)s$`


## Working with Dates
```{r}
d2 <- Sys.Date()
d2
```
### formatting dates
%d = day as number(0-31), %a = abbreviated weekday, %A = unabbreviated weekday, %m = month(00-12)
,%b = abbreviated month, %B = unabbreviated month, %y = 2 digit year, %Y = four digit year

```{r}
formate(d2, "%a %b %d")
```

### Creating dates
```{r}
x <- c("1jan1960", "2jan1960", "31mar1960", "30jul1960")
z  <- as.Date(x, "%d%b%Y")
z
z[1] - z[2]

as.numeric(z[1]- z[2])

```
### lubridate

```{r}
library(lubridate)
ymd("20140108")
mdy("08/04/2013")
dmy("03-04-2013")
```
Dealing with times
```{r}
ymd_hms("2011-08-03 10:15:03")
ymd_hms("2011-08-03 10:15:03", tz = "Pacific/Auckland")
?Sys.timezone
```

```{r}
x <- dmy(c("1jan1960", "2jan1960", "31mar1960", "30jul1960"))
wday(x[1])
wday(x[1], label = TRUE)
```
